---
title: "Data Work"
author: "K-Means Clustering"
format: 
  html:
    df-print: paged
---

Addison's Code: Libraries:

```{r, error=FALSE, warning=FALSE, message=FALSE}
library(rlang)
library(haven)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(fastDummies)
library(equatiomatic)
library(car)
library(sjPlot)
library(MASS)
library(lmtest)
library(gmodels)
library(tidyr)
library(sjmisc)
library(sjlabelled)
library(naniar)
library(cluster)
library(factoextra)
```

Read data from CSV

```{r}
#data <- read_csv("C:/Users/Addison/OneDrive/Documents/SCHOOL/Advanced Statistical Modeling/data.csv")

data <- read_csv("~/STA6257Kmeans/data.csv")
```

We know UK is the primary country associated. Making a data frame excluding UK to better see frequencies of other countries.

```{r}
clean_data <- na.omit(data)

#removing UK
clean_data_trim <- clean_data_uk[clean_data_uk$Country != "United Kingdom",]
```

Summary Stats:

```{r}
summary(clean_data)
summary(clean_data_trim)

#Plot countries
ggplot(clean_data, aes(x=Country)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))

#Create total sales column
clean_data <- clean_data %>% mutate(total = clean_data$Quantity*clean_data$UnitPrice)

#Set category breaks using quantiles
xs=quantile(clean_data$total,c(0,1/3,2/3,1))

#Label category breaks & Plot
df1 <- clean_data %>% mutate(category=cut(total, breaks=xs, labels=c("low","middle","high"),include.lowest = TRUE))
boxplot(df1$total~df1$category)

#Extra cleaning/checking
df1 <- na.omit(df1)
summary(df1)

#Creating some unique subsets for model testing
df2 <- subset(df1, Quantity > 0)
plot <- boxplot(total~category, data = df2)

#
df3 <- df2 %>% dplyr::select(Quantity,UnitPrice,CustomerID,InvoiceDate,total)
customer_data <- df3 %>% group_by(CustomerID) %>% summarize(Sales=sum(total), OrderCount=length(unique(InvoiceDate))) %>% mutate(AvgValue=Sales/OrderCount)
head(customer_data)


#Ranking and Normalization
df3_ranked <- customer_data %>% mutate(Sales=rank(Sales), OrderCount=rank(OrderCount, ties.method = "first"), AvgValue=rank(AvgValue))
df3_norm <- df3_ranked %>% mutate(Sales=scale(Sales), OrderCount=scale(OrderCount), AvgValue=scale(AvgValue))

summary(df3_norm)
sapply(df3_norm, sd)

# Create an empty vector to store WCSS values
wcss <- vector("numeric", length = 10)

# Iterate over a range of K values (e.g., from 1 to 10)
for (i in 1:10) {
  model <- kmeans(df3_norm[c("Sales", "OrderCount", "AvgValue")], centers = i, nstart = 10)
  wcss[i] <- model$tot.withinss
}

# Plot the WCSS values against the number of clusters
library(ggplot2)
ggplot(data.frame(K = 1:10, WCSS = wcss), aes(x = K, y = WCSS)) +
  geom_line() +
  geom_point() +
  labs(title = "Elbow Method to Find Optimal K", x = "Number of Clusters (K)", y = "Within-Cluster-Sum-of-Squares (WCSS)")
```

```{r}
set.seed(100)
model1 <- kmeans(df3_norm[c("Sales", "OrderCount", "AvgValue")],4)

model1

df3_norm$Cluster <- model1$cluster
```

```{r}
#Plotting Clustering Results

options(repr.plot.width = 20, repr.plot.height = 15)
ggplot(df3_norm, aes(x=AvgValue, y=OrderCount, color=Cluster)) + 
  geom_point() + 
  theme(legend.position = "right", legend.text=element_text(size=15), axis.title = element_text(size = 10))


fviz_cluster(model1, data = df3_norm,
             palette = c("blue","green","orange","red"),
             geom = "point",
             ellipse.type = "convex",
             ggtheme = theme_bw()
)
```

Aleshas code: -Guys sorry it might be a mess I will be using this space for draft work while figuring this out, and at the end ill cut out what is not needed.

```{r}
library(readr)
library(tidyverse)
library(ggplot2)
#install.packages("gtsummary")
#library(gtsummary)

data <- read_csv("~/STA6257Kmeans/data.csv")
#remove missing values
clean_data <- na.omit(data)
#removing UK
clean_data_trim <- clean_data_uk[clean_data_uk$Country != "United Kingdom",]
#View(data)
head(clean_data_trim, n=10)
names(clean_data_trim)

#tbl_summary(clean_data)

#descriptive statistics, distribution, frequencies

ggplot(clean_data, aes(x=Country)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
#Create total sales column
clean_data <- clean_data %>% mutate(total = clean_data$Quantity*clean_data$UnitPrice)

#create a data table to show example of description
clean_data %>% count(Description)

clean_data %>%top_n(5, CustomerID)

pricePlot<-ggplot(clean_data, aes(x=UnitPrice, fill=UnitPrice))+
  geom_bar(stat="identity")+
scale_fill_manual(values = c("#9900C6", "#E7B800","#3399CC","#00CC66","#FF99FF")) 
pricePlot

quantPlot<-
  ggplot(clean_data,aes(x=Quantity, color=CustomerID))+
  geom_bar()+
  scale_fill_manual(values = c("#9900C6", "#E7B800","#3399CC","#00CC66","#FF99FF"))
quantPlot

par(mfrow=c(2,2))
```

Columns: InvoiceNo, StockCode, Description, Quantity, InvoiceDate, UnitPrice, CustomerID, Country


Goals of clustering: unsupervised clustering is more concerned with classification than deep insights. Yet shines a light on features of the data that can be further discovered with deeper analytics. Explorium 2023

The goal of our analysis is to use unsupervised learning reveal patterns and relationships within this ecommerce data. This will create a deeper understanding for the entire customer-ecommerce interaction and improve customer access and experiences. While K-Means clustering creates organization where there are large amounts of unlabeled data, this analysis is just a stepping stone for clarity and deeper investigations.
