---
title: "Data Work"
author: "K-Means Clustering"
format: 
  html:
    df-print: paged
---

```{r}
library(readr)
library(ggplot2)
data <- read_csv("C:/Users/Addison/OneDrive/Documents/SCHOOL/Advanced Statistical Modeling/data.csv")
```

We know UK is the primary country associated. Making a data frame excluding UK to better see frequencies of other countries.

```{r}

clean_data_uk <- na.omit(data)

#removing UK
clean_data <- clean_data_uk[clean_data_uk$Country != "United Kingdom",]
```

Summary Stats:

```{r}
summary(clean_data_uk)
summary(clean_data)

#Plot countries
ggplot(clean_data, aes(x=Country)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle=45, vjust=1, hjust=1))
```


Aleshas code:
-Guys sorry it might be a mess I will be using this space for draft work while figuring this out, and at the end ill cut out what is not needed. 
```{r}
library(readr)
library(tidyverse)
library(ggplot2)
dataEcom <- read_csv("data.csv")
dataEcom <- dataEcom%>% filter(dataEcom$Country !="United Kingdom") %>% drop_na()
#View(data)
head(dataEcom)
str(dataEcom)
summary(dataEcom)

ggplot(dataEcom, aes(x=Country)) +
  geom_bar()+
  theme(axis.text.x =element_text(angle=45,vjust=1,hjust=1))

#Getting the amount of distinct CustomerID values and amount of distinct Descriptions
n_distinct(dataEcom$CustomerID)
n_distinct(dataEcom$Description)

#install.packages("naniar")
library(naniar)
# Plotting percentage of missing values per feature
gg_miss_var(dataEcom, show_pct = TRUE)

library(ggplot2)
#remove InvoiceNo because it does not give any specific information other than identifies/labels a transaction.
dfEcom <- as_tibble(dataEcom %>% dplyr::select(StockCode, Description,Quantity, InvoiceDate, UnitPrice, CustomerID, Country))

#how often each customer visits
countCustomerID <- dfEcom %>% mutate %>% count(as.factor(CustomerID))
countCustomerID

#transactions per country
countCountry <- dfEcom %>% mutate %>% count(as.factor(Country))
countCountry

p1<- ggplot(dataEcom, aes(x=StockCode)) +
  geom_bar()
p1

p2<- ggplot(dataEcom, aes(x=CustomerID)) +
  geom_bar()
p2


```

Columns: InvoiceNo, StockCode, Description, Quantity, InvoiceDate, UnitPrice, CustomerID, Country


Goals of clustering: unsupervised clustering is more concerned with classification than deep insights. Yet shines a light on features of the data that can be further discovered with deeper analytics. Explorium 2023

The goal of our analysis is to use unsupervised learning reveal patterns and relationships within this ecommerce data. This will create a deeper understanding for the entire customer-ecommerce interaction and improve customer access and experiences. While K-Means clustering creates organization where there are large amounts of unlabeled data, this analysis is just a stepping stone for clarity and deeper investigations.
